---
title: "ST501FinalProject_2"
author: "SpikeWang"
date: '2024-11-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2

### 2.c

The sharpe ratio is $\frac{\mu(\omega)}{\sigma(\omega)}$.

$\mu(\omega) = \omega \mu_1+(1-\omega)\mu_2$\
$\sigma(\omega) = \sqrt{\sigma_1^2\omega^2+\sigma_2^2(1-\omega)^2+2\rho\sigma_1\sigma_2\omega(1-\omega)}$\
$SR(\omega) = \frac{\omega \mu_1+(1-\omega)\mu_2}{\sqrt{\sigma_1^2\omega^2+\sigma_2^2(1-\omega)^2+2\rho\sigma_1\sigma_2\omega(1-\omega)}}$\
$\frac{d(SR(\omega))}{d\omega} = 1$\
$\frac{d^2SR(\omega)}{d\omega^2} = 1$\
When $\frac{d(SR(\omega))}{d\omega} = 0$ and$\frac{d^2SR(\omega)}{d\omega^2} \leq 0\text{, }\omega$ is the optimal value that maximize the sharpe ratio.

```{r}
mu1 <- 0.056   
mu2 <- 0.04   
sigma1 <- 0.1   
sigma2 <- 0.08 
rho <- 0.2  # correlation coefficient of 1 and 2

# calculate parameters
c0 <- mu2
c1 <- mu1 - mu2
a <- sigma1^2 + sigma2^2 - 2 * rho * sigma1 * sigma2
b <- sigma2^2 - rho * sigma1 * sigma2
c <- sigma2^2

# Sharpe Ratio function
SR <- function(w) {
  mu_w <- c0 + c1 * w
  sigma_w <- sqrt(a * w^2 - 2 * b * w + c)
  return(mu_w / sigma_w)
}

# generate weight array and Sharpe Ratio array
w_values <- seq(0, 1, length.out = 500)
sr_values <- sapply(w_values, SR)

# plot Sharpe Ratio v.s. w
plot(w_values, sr_values, type = "l", lwd = 2, col = "blue",
     main = "Sharpe Ratio SR(w) vs Portfolio Weight w",
     xlab = "Portfolio Weight w", ylab = "Sharpe Ratio SR(w)")
abline(h = 0, col = "black", lty = 2)
grid()
```

## 3

### 3.a
```{r}
# 
if (!requireNamespace("quantmod", quietly = TRUE)) install.packages("quantmod")
library(quantmod)

# download data
getSymbols(c("AAPL", "MSFT"), from = "2023-11-01", to = "2024-10-31")

# calculate daily return
AAPL.returns <- as.numeric(dailyReturn(AAPL)) * 100
MSFT.returns <- as.numeric(dailyReturn(MSFT)) * 100

# estimate moment
mu1 <- mean(AAPL.returns)
mu2 <- mean(MSFT.returns)
sigma1 <- sd(AAPL.returns)
sigma2 <- sd(MSFT.returns)
rho <- cor(AAPL.returns, MSFT.returns)

# print moment estimates
cat(sprintf("µ1 = %.4f, µ2 = %.4f, σ1 = %.4f, σ2 = %.4f, ρ = %.4f\n", mu1, mu2, sigma1, sigma2, rho))

# calculate parameter coefficient in 2(a)
c0 <- mu2
c1 <- mu1 - mu2
a <- sigma1^2 + sigma2^2 - 2 * rho * sigma1 * sigma2
b <- sigma2^2 - rho * sigma1 * sigma2
c <- sigma2^2

# print parameter coefficient in 2(a)
cat(sprintf("c0 = %.4f, c1 = %.4f, a = %.4f, b = %.4f, c = %.4f\n", c0, c1, a, b, c))
```

### 3.b

```{r}
sigma2 <- function(w) {
  a * w^2 - 2 * b * w + c
}

# generate weight w array and sigma array
w_values <- seq(0, 1, length.out = 1000)
sigma2_values <- sapply(w_values, sigma2)

# plot sigma v.s. w
plot(w_values, sigma2_values, type = "l", col = "blue", lwd = 2,
     main = expression(sigma^2(w)~vs~w),
     xlab = "w", ylab = expression(sigma^2(w)))

# find minimum sigma
opt_w <- optimize(sigma2, interval = c(0, 1))$minimum
min_sigma2 <- sigma2(opt_w)
mu_opt <- c0 + c1 * opt_w
sigma_opt <- sqrt(min_sigma2)

# output w minimizing sigma
cat(sprintf("Optimal w = %.2f, µ(w^) = %.4f, σ(w^) = %.4f\n", opt_w, mu_opt, sigma_opt))
```

### 3.c

```{r}
# calculate sharpe ratio function
sharpe_ratio <- function(w) {
  mu_w <- c0 + c1 * w
  sigma_w <- sqrt(a * w^2 - 2 * b * w + c)
  mu_w / sigma_w
}

# calculate Sharpe Ratio
sr_values <- sapply(w_values, sharpe_ratio)

# plot w v.s. sharpe ratio
plot(w_values, sr_values, type = "l", col = "blue", lwd = 2,
     main = "Sharpe Ratio SR(w) vs w",
     xlab = "w", ylab = "Sharpe Ratio SR(w)")

# optimize w to maximize Sharpe ratio
opt_sr <- optimize(sharpe_ratio, interval = c(0, 1), maximum = TRUE)
opt_w_sr <- opt_sr$maximum
max_sr <- opt_sr$objective
mu_opt_sr <- c0 + c1 * opt_w_sr
sigma_opt_sr <- sqrt(sigma2(opt_w_sr))

# print optimal w and sharpe ratio
cat(sprintf("Optimal w for Sharpe Ratio = %.2f, µ(w^) = %.4f, σ(w^) = %.4f, SR(w^) = %.4f\n",
            opt_w_sr, mu_opt_sr, sigma_opt_sr, max_sr))
```

### 3.d

```{r}
# calculate VaR function 
VaR <- function(returns, alpha) {
  quantile(-returns, probs = alpha)
}

# calculating VaR
alphas <- c(0.95, 0.975, 0.99)
VaR_AAPL <- sapply(alphas, VaR, returns = AAPL.returns)
VaR_MSFT <- sapply(alphas, VaR, returns = MSFT.returns)

# print VaR
cat("VaR for AAPL:\n")
print(round(VaR_AAPL, 4))
cat("VaR for MSFT:\n")
print(round(VaR_MSFT, 4))
```

### 3.e

```{r}
# function to calculate Portfolio VaR
VaR_portfolio <- function(w, alpha) {
  portfolio_returns <- w * AAPL.returns + (1 - w) * MSFT.returns
  VaR(portfolio_returns, alpha)
}

# calculating w and VaR(w)
for (alpha in alphas) {
  VaR_values <- sapply(w_values, VaR_portfolio, alpha = alpha)
  plot(w_values, VaR_values, type = "l", lwd = 2,
       main = bquote(VaR[.(alpha)](w)~vs~w),
       xlab = "w", ylab = bquote(VaR[.(alpha)](w)))
  
  # optimize w to minimize VaR
  opt_VaR <- optimize(VaR_portfolio, interval = c(0, 1), alpha = alpha)
  opt_w_VaR <- opt_VaR$minimum
  mu_opt_VaR <- c0 + c1 * opt_w_VaR
  sigma_opt_VaR <- sqrt(sigma2(opt_w_VaR))
  
  cat(sprintf("Optimal w for VaR (alpha=%.3f): w = %.2f, µ(w^) = %.4f, σ(w^) = %.4f\n",
              alpha, opt_w_VaR, mu_opt_VaR, sigma_opt_VaR))
}
```
